---
title: "Hands-on Exercise 2"
author: "Hao Liu"
format: html
---

```{=html}
<style>
.plot-card {
  background: #fff;
  border-radius: 8px;
  box-shadow: 0 2px 5px rgba(0,0,0,0.1);
  padding: 1rem;
  margin-bottom: 2rem;
}
</style>
```

## Get Started

Install and load required packages:

```{r}
pacman::p_load(tidyverse, RColorBrewer)
```

------------------------------------------------------------------------

## Import Data

```{r}
exam_data <- read_csv("Exam_data.csv") %>%
  mutate(Total = ENGLISH + MATHS + SCIENCE)
head(exam_data)
```

------------------------------------------------------------------------

## Visualizations

::::::: columns
:::: {.column width="50%"}
::: plot-card
**Histogram: Distribution of Total Scores**

```{r}
ggplot(exam_data, aes(x = Total)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "white", alpha = 0.7) +
  geom_density(aes(y = ..count..), color = "darkblue", size = 1) +
  theme_minimal(base_size = 14) +
  labs(title = "Distribution of Total Scores", x = "Total Score", y = "Count") +
  theme(panel.grid.minor = element_blank())
```
:::
:::::::

:::: {.column width="50%"}
::: plot-card
**Bar Chart: Average Total Score by Class**

```{r}
avg_class <- exam_data %>%
  group_by(CLASS) %>%
  summarise(Average = mean(Total)) %>%
  arrange(CLASS)

ggplot(avg_class, aes(x = CLASS, y = Average, fill = CLASS)) +
  geom_col(show.legend = FALSE) +
  theme_minimal(base_size = 14) +
  labs(title = "Average Total Score by Class", x = "Class", y = "Average Total Score") +
  theme(panel.grid.minor = element_blank(), axis.text.x = element_text(angle = 0))
```
:::
:::::::

:::: columns
:::: {.column width="50%"}
::: plot-card
**Heatmap: Correlation of Subject Scores**

```{r}
corr_mat <- exam_data %>%
  select(ENGLISH, MATHS, SCIENCE) %>%
  cor()
corr_long <- as.data.frame(as.table(corr_mat))

ggplot(corr_long, aes(x = Var1, y = Var2, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = round(Freq, 2)), size = 4) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal(base_size = 14) +
  labs(title = "Correlation Matrix of Subject Scores", x = NULL, y = NULL) +
  theme(panel.grid = element_blank())
```
:::
::::

:::: {.column width="50%"}
::: plot-card
**Bar Chart: Average Total Score by Race**

```{r}
avg_race <- exam_data %>%
  group_by(RACE) %>%
  summarise(Average = mean(Total)) %>%
  arrange(Average)

ggplot(avg_race, aes(x = reorder(RACE, Average), y = Average, fill = RACE)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  theme_minimal(base_size = 14) +
  labs(title = "Average Total Score by Race", x = NULL, y = "Average Total Score") +
  theme(panel.grid.minor = element_blank())
```
:::
::::

:::: columns
:::: {.column width="50%"}
::: plot-card
Heatmap: Average Total Score by Race and Gender

avg_rg <- exam_data %>%
  group_by(RACE, GENDER) %>%
  summarise(Average = mean(Total), .groups = 'drop')

# Average Total Score by Race and Gender
```{r}
# Calculate average total score for each Race-Gender combination
avg_rg <- exam_data %>%
  group_by(RACE, GENDER) %>%
  summarise(Average = mean(Total), .groups = 'drop')

# Create a grouped bar chart
ggplot(avg_rg, aes(x = RACE, y = Average, fill = GENDER)) +
  geom_col(position = position_dodge(width = 0.8)) +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Average Total Score by Race and Gender",
    x = "Race",
    y = "Average Total Score",
    fill = "Gender"
  ) +
  theme(panel.grid.minor = element_blank())
```
:::
::::

------------------------------------------------------------------------

© 2025 Hao Liu

