---
title: "In_class_Exercise 3"
author: Hao Liu
date: "17 May, 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
  freeze: true
---

# Loading necessary packages:

```{r library-loading, message=FALSE, warning=FALSE}
# Load libraries
pacman::p_load(tidyverse,jsonlite,
               SmartEDA,tidygraph,
               ggraph)
```

# Use fromJSON to import data:

```{r import, message=FALSE, warning=FALSE}
kg <- fromJSON("MC1_graph.json")
```

# Inspect structure:

```{r inspect, message=FALSE, warning=FALSE}
str(kg, max.level=1)
```

# Extract and Inspect:

```{r extract , message=FALSE, warning=FALSE}
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links)
```

# Initial EDA:

```{r EDA , message=FALSE, warning=FALSE}
ggplot(data = edges_tbl,
       aes(y = `Edge Type`)) +
  geom_bar()
```

# Creating Knowledge Graph

## step 1: mapping from node id to row index

```{r id_map, message=FALSE, warning=FALSE}
id_map <- tibble(id = nodes_tbl$id,
                 index = seq_len(
                   nrow(nodes_tbl)))
```

## step 2: Map source and target ids to row indices

```{r 1, message=FALSE, warning=FALSE}
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%
  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%
  rename(to = index)
```

## step 3: Creating the Graph
```{r 2, message=FALSE, warning=FALSE}
graph <- tbl_graph(nodes = nodes_tbl,
                   edges = edges_tbl,
                   directed = kg$directed)
```

## step 4: Visualising the Knowledge Graph
```{r 3, message=FALSE, warning=FALSE}
set.seed(1234)
```

```{r 4, message=FALSE, warning=FALSE}
ggraph(graph, layout = "fr") +
  geom_edge_link(alpha = 0.3, colour = "gray") +
  geom_node_point(aes(color = `Node Type`), size = 4) +
  geom_node_text(aes(label = name),
                 repel = TRUE,      # R’s logical constant
                 size = 2.5) +
  theme_void()
```

# Cleaning up

## Filter Edges to only “MemberOf”

```{r 5, message=FALSE, warning=FALSE}
graph_memberof <- graph %>%
  activate(edges) %>%
  filter(`Edge Type` == "MemberOf")
```
## Extract only connected nodes (i.e., used in these edges)
```{r 6, message=FALSE, warning=FALSE}
used_node_indices <- graph_memberof %>%
  activate(edges) %>%
  as_tibble() %>%
  select(from, to) %>%
  unlist() %>%
  unique
```
## Keep only those nodes
```{r 7, message=FALSE, warning=FALSE}
graph_memberof <- graph_memberof %>%
  activate(nodes) %>%
  mutate(row_id = row_number()) %>%
  filter(row_id %in% used_node_indices) %>%
  select(-row_id) #optional cleanup
```
## Plot the subgraph
```{r 8, message=FALSE, warning=FALSE}

ggraph(graph_memberof,
       layout = "fr") + 
  geom_edge_link(alpha = 0.5,
                 colour = "gray") +
  geom_node_point(aes(color = `Node Type`),
                  size = 1) + 
  geom_node_text(aes(label = name),
                 repel = TRUE,
                 size = 2.5) +
  theme_void()
```
